Sexual selection and speciation: a macro-evolutionary approach

### 2018-10-02 Sexual Selection and Extinction.Rmd
Critical Analysis and index add

Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM. The following is not inherently necessary to understandfor the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.

@@

```{r, fig.height = 8, fig.width = 4, warning=FALSE}
BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')

@@

BAMM.lambda.pgls.summary.RO <- dcast(BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
BAMM.lambda.pgls.summary.RO$Estimate <- NULL
BAMM.lambda.pgls.summary.RO <- sapply(BAMM.lambda.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
BAMM.lambda.pgls.summary.RO <-melt(BAMM.lambda.pgls.summary.RO) %>% na.omit()
BAMM.lambda.pgls.summary.RO$Var1 <- NULL
colnames(BAMM.lambda.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

#Get rid of outliers
BAMM.mu.pgls.summary.RO <- dcast(BAMM.mu.pgls.summary %>% select(Parameter,Estimate) %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
BAMM.mu.pgls.summary.RO$Estimate <- NULL
BAMM.mu.pgls.summary.RO <- sapply(BAMM.mu.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
BAMM.mu.pgls.summary.RO <-melt(BAMM.mu.pgls.summary.RO) %>% na.omit()
BAMM.mu.pgls.summary.RO$Var1 <- NULL
colnames(BAMM.mu.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

R-squared for relationship between RGB and Spectrophotometry Data:
```{r}
summary(lm(SDi ~ Colour.discriminability,
   data = MCC.Armenta.model.data))$r.squared %>% pander()

cor(MCC.Armenta.model.data$SDi, MCC.Armenta.model.data$Colour.discriminability)
```

@@

SS.subset %>% ggplot(aes(x = SDi, y = Sexual_selection_ppca))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_minimal()
```
**Figure S11:** The relationship between absolute sexual dichromatism and the sexual selection. Although positively correlated the distribution is very noisy with species having high sexual dichromatism and low sexual selection and visa versa.Colour.discriminability
```{r}
# summary(lm(SDi ~ Sexual_selection_ppcaColour.discriminability
#    data = SS.subset))$r.squared %>% pander()

cor(SS.subset$SDi, SS.subset$Sexual_selection_ppca)
```

@@

Subset.Lambda.pgls.summary.RO <- dcast(Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
Subset.Lambda.pgls.summary.RO$Estimate <- NULL
Subset.Lambda.pgls.summary.RO <- sapply(Subset.Lambda.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
Subset.Lambda.pgls.summary.RO <-melt(Subset.Lambda.pgls.summary.RO) %>% na.omit()
Subset.Lambda.pgls.summary.RO$Var1 <- NULL
colnames(Subset.Lambda.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

Subset.Mu.pgls.summary.RO <- dcast(Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter)
Subset.Mu.pgls.summary.RO$Estimate <- NULL
Subset.Mu.pgls.summary.RO <- sapply(Subset.Mu.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
Subset.Mu.pgls.summary.RO <-melt(Subset.Mu.pgls.summary.RO) %>% na.omit()
Subset.Mu.pgls.summary.RO$Var1 <- NULL
colnames(Subset.Mu.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

BAMM.Lambda.100Trees %>% ggplot(aes(x = mean.lambda, y = DR.Lambda.100Trees$DR))+
  geom_point()+
  geom_errorbarh(aes(xmin = mean.lambda - 1.96*sd, xmax = mean.lambda + 1.96*sd), size = 0.05)+
  geom_errorbar(aes(ymin = DR.Lambda.100Trees$DR - 1.96*DR.Lambda.100Trees$sd, ymax = DR.Lambda.100Trees$DR + 1.96*DR.Lambda.100Trees$sd), size = 0.05)+
  theme_minimal()

# mean(DR.Lambda.100Trees$sd)
# mean(BAMM.Lambda.100Trees$sd)
```
**Figure S14:**

```{r, eval = FALSE, echo = F}
all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Rates.100Trees %>% select(mean.lambda, TipLabel))
cor(all.trees.join$DR, all.trees.join$mean.lambda)
```

```{r}
MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')
restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda))+
  geom_point()+
  theme_minimal()
```
**Figure S15:**

```{r, eval = FALSE, echo = F}
MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel))
cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda)
```

@@

What if we run the subset df qwith SDi
```{r, eval = FALSE, echo = F}
#Run model for DR
Subset.MCC.DR.SDi <- gls(MCC.DR ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
                data = SS.subset, 
                method = "REML")

>>>>

Given the high variability in tip-rate estimates from the above plot, below we peform some diagnostics on BAMM to demonstrate that the variability is unlikely an error in sampling or parameters, rather an inherent aspect of BAMM. The following is not inherently necessary to understand the conclusions drawn from the paper, however they do raise a set of methodological questions about BAMM that warrant further investigation.

@@

```{r, warning=F, error = F, fig.height = 8, fig.width = 4}
BAMM.lambda.pgls.models <- readRDS("data/BAMM.lambda.pgls.models.rds")
MCC.Lambda.top <- readRDS('data/MCC.Lambda.top.rds')

@@

BAMM.lambda.pgls.summary.RO <- dcast(BAMM.lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
BAMM.lambda.pgls.summary.RO$Estimate <- NULL
BAMM.lambda.pgls.summary.RO <- sapply(BAMM.lambda.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
BAMM.lambda.pgls.summary.RO <-melt(BAMM.lambda.pgls.summary.RO) %>% na.omit()
BAMM.lambda.pgls.summary.RO$Var1 <- NULL
colnames(BAMM.lambda.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

#Get rid of outliers
BAMM.mu.pgls.summary.RO <- dcast(BAMM.mu.pgls.summary %>% select(Parameter,Estimate) %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
BAMM.mu.pgls.summary.RO$Estimate <- NULL
BAMM.mu.pgls.summary.RO <- sapply(BAMM.mu.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
BAMM.mu.pgls.summary.RO <-melt(BAMM.mu.pgls.summary.RO) %>% na.omit()
BAMM.mu.pgls.summary.RO$Var1 <- NULL
colnames(BAMM.mu.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

R-squared and correlation for relationship between RGB and Spectrophotometry Data:
```{r}
data_frame(
R2 = summary(lm(SDi ~ Colour.discriminability,
   data = MCC.Armenta.model.data))$r.squared,
r = cor(MCC.Armenta.model.data$SDi, MCC.Armenta.model.data$Colour.discriminability)) %>% pander()
```

@@

SS.subset %>% ggplot(aes(x = SDi, y = Sexual_selection_ppca))+
  geom_point()+
  geom_smooth(method = "lm")+
  theme_minimal()
```
**Figure S11:** The relationship between absolute sexual dichromatism and the sexual selection. Although positively correlated the distribution is very noisy with species having high sexual dichromatism and low sexual selection and visa versa.Colour.discriminability
```{r}
data_frame(
R2 = summary(lm(SDi ~ Sexual_selection_ppca,
   data = SS.subset))$r.squared,
r = cor(SS.subset$SDi, SS.subset$Sexual_selection_ppca)) %>% pander(digits = 2)
```

@@

Subset.Lambda.pgls.summary.RO <- dcast(Subset.Lambda.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
Subset.Lambda.pgls.summary.RO$Estimate <- NULL
Subset.Lambda.pgls.summary.RO <- sapply(Subset.Lambda.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
Subset.Lambda.pgls.summary.RO <-melt(Subset.Lambda.pgls.summary.RO) %>% na.omit()
Subset.Lambda.pgls.summary.RO$Var1 <- NULL
colnames(Subset.Lambda.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

Subset.Mu.pgls.summary.RO <- dcast(Subset.Mu.pgls.summary %>% filter(Parameter != "(Intercept)"), Estimate ~ Parameter, value.var = "Estimate")
Subset.Mu.pgls.summary.RO$Estimate <- NULL
Subset.Mu.pgls.summary.RO <- sapply(Subset.Mu.pgls.summary.RO, function(x) {
  remove_outliers(x, na.rm = T)})
Subset.Mu.pgls.summary.RO <-melt(Subset.Mu.pgls.summary.RO) %>% na.omit()
Subset.Mu.pgls.summary.RO$Var1 <- NULL
colnames(Subset.Mu.pgls.summary.RO) <- c("Parameter", "Estimate")

@@

all.trees.join <- left_join(DR.Lambda.100Trees %>% select(TipLabel, DR), BAMM.Lambda.100Trees %>% select(mean.lambda, TipLabel), by = "TipLabel")

BAMM.Lambda.100Trees %>% ggplot(aes(x = DR.Lambda.100Trees$DR, y = mean.lambda, fill = mean.lambda))+
  geom_point(shape = 21)+
  geom_errorbar(aes(ymin = mean.lambda - 1.96*sd, ymax = mean.lambda + 1.96*sd), size = 0.02)+
  geom_errorbarh(aes(xmin = DR.Lambda.100Trees$DR - 1.96*DR.Lambda.100Trees$sd, xmax = DR.Lambda.100Trees$DR + 1.96*DR.Lambda.100Trees$sd), size = 0.02)+
  theme_minimal()+
  ylab("100 Trees Speciation Rate (BAMM)")+
  xlab("100 Trees Speciation Rate (DR)")+
  scale_fill_viridis_c()

# mean(DR.Lambda.100Trees$sd)
# mean(BAMM.Lambda.100Trees$sd)
```
**Figure S14:** Speciation Rate means from 100 trees using either the DR statistic (x axis) or BAMM (y axis). While there is a clear correlation (*r* = `r round(cor(all.trees.join$DR, all.trees.join$mean.lambda), digits = 2)`) there is variability with BAMM showing less heterogeneity. 95 % CIs are plotted for both axis from the 100 trees. Each point represents a species (*n* = 5,965).

```{r, echo = F}
MCC.BAMM.model.data <- readRDS('data/MCC.BAMM.model.data.rds')

#cor(all.trees.join$DR, all.trees.join$mean.lambda)

MCC.tree.join <- left_join(restricted.data %>% select(MCC.DR, TipLabel), MCC.BAMM.model.data %>% select(mean.lambda, TipLabel), by = "TipLabel")
#cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda)
```

```{r, fig.width = 8, fig.height = 6}
restricted.data %>% ggplot(aes(x = MCC.DR, y = MCC.BAMM.model.data$mean.lambda, fill = MCC.DR))+
  geom_point(shape = 21)+
  theme_minimal()+
  geom_smooth(method = "lm")+
  ylab("MCC Speciation Rate (BAMM)")+
  xlab("MCC Speciation Rate (DR)")+
  scale_fill_viridis_c()

```
**Figure S15:** Similar to Figure S14, there is a correlation (*r* = `r round(cor(MCC.tree.join$MCC.DR, MCC.tree.join$mean.lambda), digits = 2)`) between $\lambda_{DR}$ and $\lambda_{BAMM}$ with BAMM results showing less heterogeneity. Each point represents a species (*n* = 5,965).

@@

```{r, eval = FALSE, echo = F}
#Run subset model with SDi
#Run model for DR
Subset.MCC.DR.SDi <- gls(MCC.DR ~ SDi
                         + log(range.size.m2)
                         + bioclim4 #Seasonal variation
                         + residuals.PC1 #Spatial variation
                         + PC1.LIG #Long-term climate variation
                         + NPP,
                correlation = corPagel(1, phy = pruned.MCC.Subset.tree, fixed = FALSE), 
                data = SS.subset, 
                method = "REML")
